#Name: Sweta Kumari 
#ANALYSIS OF COVARIANCE

#ANCOVA calculations
```{r}
photoperiod<-read.table("C:\\Users\\Sweta\\Documents\\Data\\Photoperiod.txt",header=T)
attach(photoperiod)
names(photoperiod) 
```
#Let say Photoperiod=x and Growth=y
#Find the sum of x and sum of x square
```{r}
sum(Photoperiod);sum(Photoperiod^2)
```
#Find the sum of y and sum of y square
```{r}
sum(Growth);sum(Growth^2) 
```
#Find the sum of x*y
```{r}
sum(Growth*Photoperiod)
```
#Find the sum and sum of squares
#Calculate the sum of growth according to Genotype
```{r}
tapply(Growth,Genotype,sum)
```
#Calculate the sum of growth square according to Genotype
```{r}
tapply(Growth^2,Genotype,sum) 
```
#Calculate the sum of Photoperiod according to Genotype
```{r}
tapply(Photoperiod,Genotype,sum)
```
#Calculate the sum of Photoperiod square according to Genotype
```{r}
tapply(Photoperiod^2,Genotype,sum) 
```
#Calculate the sum of growth*Photoperiod according to Genotype
```{r}
tapply(Photoperiod*Growth,Genotype,sum)
```
#Analysis of Covariance in R
#Use lm directive and capture the output in model object
```{r}
options(contrasts=c("contr.treatment","contr.poly")) 
model<-lm(Growth~Genotype*Photoperiod) 
```
#Summarize the model
```{r}
summary(model)
```
#Use aov directive to get anova table for model
```{r}
model<-aov(Growth~Genotype*Photoperiod) 
```
#Summarize the model
```{r}
summary(model) 
```
#use summary.aov to get ANOVA table
```{r}
summary.aov(model)
```
#Use summary.lm to get the parameters and standard error.
```{r}
summary.lm(model) 
```
#Ancova with different values of the covariates
```{r}
ipomopsis<-read.csv("C:\\Users\\Sweta\\Documents\\Data\\ipomopsis.csv",header=T)
attach(ipomopsis)
names(ipomopsis)
```
#Find sum of root and root^2
```{r}
sum(Root);sum(Root^2)
```
#Find sum of fruit and fruit^2
```{r}
sum(Fruit);sum(Fruit^2) 
```
#Find sum of root*fruit
```{r}
sum(Root*Fruit)
```
#Let's consider only grazed plant data
```{r}
sum(Root[Grazing=="Grazed"]);sum(Root[Grazing=="Grazed"]^2) 
```
#Now consider ungrazed plant data
```{r}
sum(Root[Grazing=="Ungrazed"]);sum(Root[Grazing=="Ungrazed"]^2)
```
#Let's check data on fruits for grazed plant data
```{r}
sum(Fruit[Grazing=="Grazed"]);sum(Fruit[Grazing=="Grazed"]^2)
```
#Similarly fruit data for ungrazed plant 
```{r}
sum(Fruit[Grazing=="Ungrazed"]);sum(Fruit[Grazing=="Ungrazed"]^2) 
```
#To get data for sum of products of root and fruit for grazed plant.
```{r}
sum(Root[Grazing=="Grazed"]*Fruit[Grazing=="Grazed"]) 
```

#Similarly for ungrazed plant
```{r}
sum(Root[Grazing=="Ungrazed"]*Fruit[Grazing=="Ungrazed"])
```

#Ancova in R using lm 
#Find the mean of fruit for grazed and ungrazed plant
```{r}
tapply(Fruit,Grazing, mean)
```
#Let's plot the graph
```{r}
plot(Root,Fruit)
```

#Let's do blank plot by type=n; #select grazed and ungrazed points for fruits and root;# Use plotting character = 16 for plotting the points
```{r}
plot(Root,Fruit,type="n") 
points(Root[Grazing=="Ungrazed"],Fruit[Grazing=="Ungrazed"])
points(Root[Grazing=="Grazed"],Fruit[Grazing=="Grazed"],pch=16)
```


#To get different slopes and inercepts of both variables, use * asterisk mark .
```{r}
ancova<-lm(Fruit~Grazing*Root) 
```
#Summarize the result
```{r}
summary(ancova)
```
#ANOVA table for maximal model
```{r}
anova(ancova) 
```
#update the ancova model by removing the interaction between grazing and roots
```{r}
ancova2<-update(ancova, ~ . - Grazing:Root)
```
#Compare ancova and ancova2 model
```{r}
anova(ancova,ancova2) 
```
#Update ancova2 model by removing Grazing from the model
```{r}
ancova3<-update(ancova2, ~ . - Grazing) 
```
#Compare ancova2 and ancova3 model
```{r}
anova(ancova2,ancova3)
```
#Summary of ancova2 model
```{r}
summary(ancova2) 
```
#ANOVA table for ancova2 model
```{r}
anova(ancova2)
```
#Do model simplification using step directive called as automatic model-simplification directive
```{r}
step(ancova)
```
#Summary of ancova2
```{r}
summary(ancova2)
```
#Use abline directive to get the line; calling plot directive again
```{r}
plot(Root,Fruit)
points(Root[Grazing=="Ungrazed"],Fruit[Grazing=="Ungrazed"])
points(Root[Grazing=="Grazed"],Fruit[Grazing=="Grazed"],pch=16)
abline(-127.829 , 23.56)
abline(-127.829+36.1032 , 23.56)
```

#A more complex Ancova: 2 factors and 1 continuous covariate 
```{r}
gain<-read.table("C:\\Users\\Sweta\\Documents\\Data\\Gain.txt",header=T)
attach(gain)
names(gain) 
```
#We begin by considering all interaction
```{r}
m1<-lm(Weight~Sex*Age*Genotype)
```
#summarize m1 model
```{r}
summary(m1)
```
#Model simplication
```{r}
step(m1)
```
#Consider only variables (Don't consider interaction between the variables)
```{r}
m2<-aov(Weight~Sex+Age+Genotype)
```
#SUmmary of model m2
```{r}
summary(m2)
```
#To get standard error and other parameters
```{r}
summary.lm(m2)
```
#Factor level reduction
```{r}
newgen<-factor(1+(Genotype=="CloneB")+(Genotype=="CloneD")+ 2*(Genotype=="CloneC") + 2*(Genotype=="CloneE") + 3*(Genotype=="CloneF")) 
newgen
```
#Use Sex, Age, newgen and define new model m3
```{r}
m3<-lm(Weight~Sex+Age+newgen)
```
#Compare anova table for m2 and m3
```{r}
anova(m2,m3) 
```





