#Name: Sweta Kumari 
#ANALYSIS OF VARIANCE

#1. One-way Analysis of Variance:
#Read the oneway file.
```{r}
oneway<-read.table("C:\\Users\\Sweta\\Documents\\Data\\oneway.txt",header=T)
attach(oneway)
names(oneway) 
```
#Calculate mean of growth using tapply function:
```{r}
tapply(Growth,Photoperiod,mean)
```
#To change the order of the output in Very.short,Short,Long,Very.long
```{r}
Photoperiod<-ordered(Photoperiod,levels=c("Very.short","Short","Long","Very.long"))
```
#Caluclate mean for new ordered output:
```{r}
tapply(Growth,Photoperiod,mean) 
```
#To find Photoperiod is factor variable
```{r}
is.factor(Photoperiod) 
```
#Use aov directive to show the anova table for the "oneway.txt" dataset.Capture the result in one.way object.
```{r}
one.way<-aov(Growth~Photoperiod) 
```
#Use summary function to see the result of anova table
```{r}
summary(one.way) 
```
#2. Two way analysis of variance
#Remove Growth and Photoperiod variables and detach the oneway.txt dataset
```{r}
rm("Growth", "Photoperiod")
detach(oneway)
```
#Read the twoway file.
```{r}
twoway<-read.table("C:\\Users\\Sweta\\Documents\\Data\\twoway.txt",header=T)
attach(twoway)
names(twoway)
```
#Find the mean growth rate for 6 genotypes
```{r}
tapply(Growth,Genotype,mean) 
```
#Use aov directive to show the anova table for the "twoway.txt" dataset.Capture the result in two.way object.
```{r}
two.way<-aov(Growth~Genotype+Photoperiod) 
```
#Use summary function to see the result of anova table
```{r}
summary(two.way) 
```
#3. Two-way Factorial Analysis of Variance
#Read the factorial file.
```{r}
factorial<-read.table("C:\\Users\\Sweta\\Documents\\Data\\factorial.txt",header=T)
factorial
attach(factorial) 
```
#Use aov directive to show the anova table for the "factorial.txt" dataset.Capture the result in model object.
```{r} 
model<-aov(growth~diet*coat) 
```
#Use summary function to see the result of anova table
```{r}
summary(model)
```

#Find sum of growth for all 6 combinations
```{r}
tapply(growth,list(coat,diet),sum)
```

#Sum of squares of subtotal
```{r}
SSAB<-sum(as.vector(tapply(growth,list(coat,diet),sum))^2)/2 
```

#Correction factor
```{r}
CF<-sum(growth)^2/length(growth) 
```
#Fins sum of squares (SSAB–SSA-SSB-CF)
```{r}
SSAB-CF-2.66-2.61333 
```
#Update model and remove interaction term
```{r}
model2<-update(model ,  ~ .  -  diet:coat) 
```
#Use anova directive to compare both models.
```{r}
anova(model,model2) 
```
#Summary of variance table for model2
```{r}
summary(model2)
```
#Remove diet variable as p value is not significant.
```{r}
model3<-update(model2, ~. -diet) 
```
#Compare model3 and model2
```{r}
anova(model2,model3)
```
#Summary of model3
```{r}
summary(model3) 
```
#To find mean for all three diet
```{r}
tapply(growth,diet,mean)
```
#New factor for diet: 1 for A, B and 2 for C
```{r}
diet2<-factor(1+(diet=="C"))
diet2
```
#Update model3 and add diet2
```{r}
model4<-update(model3, ~. +diet2) 
```
#Compare model3 and model4
```{r}
anova(model3,model4) 
```
#To check for interaction between diet2 and coat
```{r}
model5<-update(model4, ~. +diet2:coat)
```
#Compare model 4 and model5
```{r}
anova(model4,model5) 
```
#model5 is inadequate so model4 is minimal adequate model
```{r}
summary(model4)
```
#3. Three-way Factorial Analysis of Variance
#Read the file
```{r}
Daphnia.data<-read.table("C:\\Users\\Sweta\\Documents\\Data\\Daphnia.txt",header=T)
attach(Daphnia.data)
names(Daphnia.data) 
```
#Use aov directive to show the anova table for the "factorial.txt" dataset.Capture the result in factorial object.
```{r}
factorial<-aov(Growth.rate~Water*Detergent*Daphnia) 
``` 
#Summary function for anova table
```{r}
summary(factorial)
```
#Create a 2dimensional table of mean growth rates
```{r}
tapply(Growth.rate,list(Detergent,Daphnia),mean) 
```
#To show the interaction effects
```{r}
interaction.plot(Detergent,Daphnia,Growth.rate) 
```

