#Name: Sweta Kumari 
#STATISTICS OF ONE AND TWO SAMPLES

#Central tendency
#1. Mode
#Let's define x 
```{r}
x<-(1:9)
rm(x)
```
#Read mode.txt from the directory.
```{r}
distribution<-read.table("C:\\Users\\Sweta\\Documents\\Data\\mode.txt", header=T)
attach(distribution)
names(distribution)
```
#Draw 2 histograms side by side: Mode gives idea about the central tendency of the data set.
```{r}
par(mfrow=c(1,2))
barplot(fx,names=as.character(x))
barplot(fy,names=as.character(x))
```

#2.Median: Another way of measuring central tendency is Median.Define y vector where it has all elements of x values.
```{r}
y<-rep(x,fx)
y
```
#Sort the value of Y. #Determine how many y values we have in y vector.
```{r}
sort(y)
length(y)
```
#Since length is odd, we can find median by determining 17th value of y.
```{r}
ceiling(length(y)/2)
```
#Find the value of the y in 17th position
```{r}
y[17]
```
#Or, we can determine y value at 17th position:
```{r}
y[ceiling(length(y)/2)]
```
#We can find the median with built-in function also:
```{r}
median(y)
```
#3. Arithmetic mean:We can calculate this by sum and length function
```{r}
sum(y)/length(y)
```
#or use built-in function
```{r}
mean(y)
```
#4. Geometric Mean:
```{r}
prod(y)^(1/length(y))
```
#another way of finding the Geometric mean is find the mean of log y and then take antilog. Here log has base e.
```{r}
meanlogy<-sum(log(y))/length(y) 
meanlogy
exp(meanlogy) 
```
#There is no built-in function for geometric mean but we can write as follows:
```{r}
geometric<-function(x) exp(sum(log(x))/length(x))
geometric(y) 
```
#Geometric mean is useful in extreme data set. Foe instance, consider aphids count in plants:
```{r}
aphids<-c(10,1,1,10,1000)
```
#We can find Arithmetic mean but it will not give measure of central tendency
```{r}
mean(aphids)
```
#Geometric mean can be found to know more about the central tendency of the data
```{r}
geometric(aphids)
```
#5. Harmonic Mean:Let's find mean of the all 4 velocities
```{r}
mean(c(1,2,4,1))
```
#To find the Harmonic mean:
```{r}
4/sum(1/c(1,2,4,1))
```
#Let's write harmonic mean formula:
```{r}
harmonic<-function(x) 1/mean(1/x)
harmonic(y)
```
#mode(y) will not give mode of the dataset. It suggests the type of the y (numeric).
```{r}
mode(y)
```
#Frequnecy distribution of fz:
```{r}
par(mfrow=c(1,1)) 
barplot(fz,names=as.character(x))
```

#Define w and find all information for central tendency:Here largets mode is at x=11.
```{r}
w<-rep(x,fz)
mean(w)
median(w)
geometric(w)
harmonic(w)
```
#6. Measuring variation: Define y as follows and plot the graph:
```{r}
y<-c(13,7,5,12,9,15,6,11,9,7,12)
plot(y,ylim=c(0,20)) 
```

#To find the variability of the y, find range of y:
```{r}
range(y)
```
#To plot the line with average value of y and zero slope :
```{r}
y<-c(13,7,5,12,9,15,6,11,9,7,12)
plot(y,ylim=c(0,20))
abline(mean(y),0)
```

#7. Find the variance: Define A,B,C and find mean for all vectors
```{r}
A<-c(3,4,4,3,2,3,1,3,5,2)
B<-c(5,5,6,7,4,4,3,5,6,5)
C<-c(3,3,2,1,10,4,3,11,3,10)
mean(A)
mean(B)
mean(C)
```
#Find the differences y- ybar and then find the sum of suquares
```{r}
dA<-A-3
dB<-B-5
dC<-C-5
SSA<-sum(dA^2)
SSB<-sum(dB^2)
SSC<-sum(dC^2)
```
#Find the variance by dividing sum of squares by degree of freedom
```{r}
s2A<-SSA/9
s2B<-SSB/9
s2C<-SSC/9 
s2A;s2B;s2C
```
#Bulit-in function to find variance
```{r}
s2A<-var(A)
s2A
```
#To know if variance of garden c is significantly higher than other variance. #Variance of garden C is significantly higher as it is more than 4 times the variance of other garden.
```{r}
14.2222 / 1.3333
```

#Standard error of market gardens A, B, C
```{r}
sqrt(s2A/10)
```
```{r}
sqrt(s2B/10)
```
```{r}
sqrt(s2C/10)
```
#8. Find the t value and confidence interval:
#Find t value at 95% Confidence interval for two tail test; t-value at left tail:
```{r}
qt(.025,9)
```
#Find the t value for right tail:
```{r}
qt(.975,9)
```
#Find t value at 99% Confidence interval for two tail test
```{r}
qt(.995,9)
```
#Find t value at 99.5% Confidence interval for two tail test
```{r}
qt(.9975,9)
```
#Confidence interval for garden B
```{r}
qt(.975,9)*sqrt(1.33333/10)
```
#9. Quantiles:Consider z with 1000 random numbers from standard normal distribution where mean is 0 and standard deviation is 1.
```{r}
z<-rnorm(1000)
```
#calculate mean
```{r}
mean(z)
```
#calculate quantiles
```{r}
quantile(z,c(.025,.975))
```
#consider z with 10000 random numbers from standard normal distribution and find quantiles.
```{r}
z<-rnorm(10000)
quantile(z,c(.025,.975))
```
#10. Robust estimators:Define y as follows 
```{r}
y<-c(3,4,6,4,5,2,4,5,1,5,4,6)
```
#Find the Median Absolute Deviation
```{r}
mad(y)
```
#Find the standard deviation
```{r}
sd(y)
```
#Define new vector y1 by adding 100 as outleir in the existing vector y
```{r}
y1<-c(y,100)
```
#find the mean of y1
```{r}
mean(y1)
```
#find the standard deviation of y1
```{r}
sqrt(var(y1))
```
#Find the Median Absolute Deviation of y1
```{r}
mad(y1)
```
#Create function outlier:If standard deviation is greater than 4 times Median Absolute Deviation then it is outlier.
```{r}
outlier<-function(x) {
  if(sqrt(var(x))>4*mad(x))   print("Outliers present")     
  else   print("Deviation reasonable")}
```
```{r}
#Check for vector y
outlier(y)
```
```{r}
#Check for vector y1
outlier(y1) 
```
#11. Single sample estimation:
#Read the light file
```{r}
light<-read.table("C:\\Users\\Sweta\\Documents\\Data\\light.txt",header=T) 
attach(light)
names(light)
```
#Plot histogram for speed
```{r}
hist(speed)
```

#summary of speed
```{r}
summary(speed)
```
#12. Inference in 1-sample case:
#Conduct Wilcoxon’s signed rank test for speed of light.
```{r}
library(stats)
wilcox.test(speed,mu=990)
```
#Conduct student's t-test.
```{r}
t.test(speed,mu=990)
```
#13. Comparing two means: find the critical value of student's t
```{r}
qt(.975,18)
```
#Value of t-statistics
```{r}
(mean(A)-mean(B))/sqrt(s2A/10+s2B/10)
```
#We can use built-in function to calculate the t-statistics
```{r}
t.test(A,B) 
```
#14. Wilcoxon rank sum test: Plot the histogram for garden B and C to check the non-normality of the error:
```{r}
par(mfrow=c(1,2))
hist(B,breaks=c(0.5:11.5))
hist(C,breaks=c(0.5:11.5))
```

#Combine both vectors B and C
```{r}
combined<-c(B,C)
combined
```
#Create a list with sample name B and C
```{r}
sample<-c(rep("B",10),rep("C",10))
sample
```
#Find the rank of the combined vector
```{r}
rank.combi<-rank(combined)
rank.combi
```
#Calculate the sum of the ranks for garden B.
```{r}
sum(rank.combi[sample=="B"])
```
#Calculate the sum of the ranks for garden C.
```{r}
sum(rank.combi[sample=="C"])
```
#Or,Use tapply function to calculate the sum of the ranks for garden B and C.
```{r}
tapply(rank.combi,sample,sum)
```
#Conduct Wilcoxon rank sum test with built-in function
```{r}
wilcox.test(B,C)
```
#15. Tests on paired samples: Define x and y as follows
```{r}
x<-c(20,15,10,5,20,15,10,5,20,15,10,5,20,15,10,5)
y<-c(23,16,10,4,22,15,12,7,21,16,11,5,22,14,10,6)
```
#Conduct student's t-test:
```{r}
t.test(x,y)
```
#Conduct student's paired t-test:
```{r}
t.test(x,y,paired=T)
```
#16. Central Limit Theorem : Define y with 1000 random number of negative binomial random numbers with parameters size as 1, probability as 0.2
```{r}
par(mfrow=c(1,2))
y<-rnbinom(1000,1,.2)
```
#Calculate the mean of Y
```{r}
mean(y)
```
#Calculate the variance of Y
```{r}
var(y)
```
#To get the table for frequnecy distribution
```{r}
table(y)
```
#Plot histogram with break points.Then plot histogram from the same set of y with sample size as 30 
```{r}
par(mfrow=c(1,2))
hist(y,breaks=-0.5:38.5)
my <- numeric(1000) 
for (i in 1:1000) {
  y <- rnbinom(30, 1, 0.2)  
  my[i] <- mean(y) } 
hist(my)
```
```{r}
par(mfrow=c(1,1))
```