#Name: Sweta Kumari 
#REGRESSION

#1. Data Inspection:Read the regression file and plot the graph between tannin and growth. Find the mean of growth then plot line with slope as 0 and inetrcept as mean of growth
```{r}
regression<-read.csv("C:\\Users\\Sweta\\Documents\\Data\\regression.csv",header=T)
attach(regression)
names(regression)
plot(tannin,growth)
mean(growth)
abline(6.889,0)
```

#The quantile of Student’s t distribution (2-tailed, alpha= 0.025) with n-2 degrees of freedom.
```{r}
pairs(regression,panel = panel.smooth)
qt(.975,7)
```
#Calculations involved in linear regression
```{r}
regression
```
#Let say x is tannin and y is growth
#Find sum of x and sum of x^2
```{r}
sum(tannin);sum(tannin^2)
```
#Find sum of y and sum of y^2
```{r}
sum(growth);sum(growth^2)
```
#Find sum of xy
```{r}
sum(tannin*growth)
```
#quantiles of the F distribution
```{r}
qf(0.95,1,7)
```
#The probability of getting an F value of 30.98 if the null hypothesis of b = 0 is true.
```{r}
1-pf(30.98,1,7)
```
#To calculate the 95% confidence intervals: 2-tailed value of Student's t with d.f=7.
```{r}
qt(.975,7)
```
#Using R for regression: Obtain scatter plot then add fitted line abline directive with lm (linear model)
```{r}
plot(tannin,growth)
abline(lm(growth~tannin))
```

#Let's called model object as "model"
```{r}
model<-lm(growth~tannin)
```
#Summarize the model:
```{r}
summary(model)
```
#Plot the model with two graphs in two rows
```{r}
par(mfrow=c(2,2))
plot(model) 
```

#the predicted value of y hat  at tannin = 5.5%
```{r}
11.7556-1.2167*5.5
```
#Alternate way is to use predict directive with object called model
```{r}
predict(model,list(tannin=5.5))
```
#We will use mfrow to plot one graph in a row. If there is any x and y vector exist, it can be removed by rm:
```{r}
par(mfrow=c(1,1))
rm(x,y)
```
#Transformation of non-linear responses
```{r}
decay<-read.csv("C:\\Users\\Sweta\\Documents\\Data\\decay.csv",header=T)
attach(decay)
names(decay)
x<-time
y<-amount
plot(x,y)
abline(lm(y~x))
```

#Let's called model object as result
```{r}
result<-lm(y~x)
summary(result)
```

#Diagnostic plot
```{r}
plot(result) 
```

#Plot the graph for x vs ln y
```{r}
plot(x,y,log="y")
```

#Let's say new model object is "transformed"
```{r}
transformed <- lm(log(y)~x) 
summary(transformed) 
```

#Let's check the diagmostic plot:
```{r}
plot(transformed)
```

#Let's redraw the scatter plot for x and y
```{r}
plot(x,y)
```

#Let's generate sequence from 0 to 30 with inetrval 0.1 and predict the respective value of y:
```{r}
smoothx<-seq(0,30,0.1) 
smoothy<-exp(predict(transformed,list(x=smoothx)))
```
#Draw the fitted line:
```{r}
plot(x,y)
lines(smoothx,smoothy)
```
