#Name: Sweta Kumari
#Time series analysis
#Read the global temperture file and check the columns name.
#1. Read the global temperture file and check the columns name.
```{r}
GlobalTemperature<-read.csv("C:\\Users\\Sweta\\Documents\\Data\\GlobalTemperatures.csv",header=T)
attach(GlobalTemperature)
names(GlobalTemperature)
```

#2. Importing librarries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
library(fpp2)
library(zoo)
```

#3.#To review the changes in the 'LandAverageTemperature', we plot the graph of LandAverageTemperature vs time.
#Analysis: From the graph, there is no outlier observed in the graph. 
#The Land Average Temperature is between 0 and 16 for the period of 1850 to 2015.
#As the years passed, the average tempertaure of land has increased.
#So, Average temperature is not stationary and showing upward trend.
```{r}
Temperature <- GlobalTemperature %>% mutate(Time_year= year(dt)) %>% filter(Time_year>1849)
Temperature %>% group_by(Time_year) %>% ggplot(aes( x=Time_year,y=LandAverageTemperature,Colour=LandAverageTemperature)) + geom_line()
```
#4. To review the changes in the 'LandAndOceanAverageTemperature', we plot the graph of LandAndOceanAverageTemperature vs time.
#Analysis: From the graph, there is no outlier observed in the graph. 
#The Land and Ocean Average Temperature is between 12 and 18 for the period of 1850 to 2015.
#As the years passed, the average tempertaure has increased.
#So, Average temperature is not stationary and showing upward trend.
#We will check the seasonality with help of moving average method in last section
```{r}
Temperature %>% group_by(Time_year) %>% ggplot(aes( x=Time_year,y=LandAndOceanAverageTemperature,Colour=LandAndOceanAverageTemperature)) + geom_line()
```

#5. To review the changes in the 'LandMaxTemperature', we plot the graph of LandMaxTemperature vs time.
#Analysis: From the graph, there is no outlier observed in the graph. 
#The Land Maximum Temperature is between 6 and 21 for the period of 1850 to 2015.
#As the years passed, the maximum tempertaure of land has increased.
#So, Land Max temperature is not stationary and showing upward trend.
```{r}
Temperature %>% group_by(Time_year) %>% ggplot(aes( x=Time_year,y=LandMaxTemperature,Colour=LandMaxTemperature)) + geom_line()
```

#6.Running a 10-year moving average on LandAndOceanAverageTemperature.
#Analysis: Moving average helped to smooth the graph. Graph is showing strong seasonality with upward trend.
#In 1880 and 1950, there is increment in temperature. From 1980, there is continuous increment in the temperature and it has crossed 16 degree in 2015.
```{r}
LandTemperature <- GlobalTemperature  %>% mutate(LandAndOceanAverageTemperature = rollmean(LandAndOceanAverageTemperature, k = 120, fill = NA, align = "right"), Time_year=year(dt))
LandTemperature %>% group_by(Time_year) %>% filter(Time_year>1849) %>% ggplot(aes( x=Time_year,y=LandAndOceanAverageTemperature,Colour=LandAndOceanAverageTemperature)) + geom_line()
```



